Mapping sites of shifting and constant mutational effects on the evolutionary landscape of HIV Envelope

### 2017-10-24 paper/manuscript.tex
moved manuscript into `./paper/` subdirectory

\begin{figure}
\centerline{\includegraphics[width=5in]{figures/tree/tree}}
\caption{\label{fig:tree}
Phylogenetic tree showing the relationship among BG505 and BF520.
{\bf(A)} A tree of group-M sequences with 20 sequences per subtype for subtypes A, B, C, D, F, and G.
{\bf(B)} A tree of subtype-A sequences with 120 sequences total, rooted using a subtype-B sequence as an outgroup.
BG505 and BF520 cluster within subtype A, while LAI clusters within subtype B.
The alignments we used to make these trees are the same alignments we used in the below \texttt{phydms} analysis, excluding the outgroup sequence used to root the subtype-A tree.
The scale bars are in units of codon substitutions per site.
}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=5in]{figures/mutfreqs/mutfreqs}}
\caption{\label{fig:mutfreqs}
The deep mutational scanning experiments imposed strong purifying selection and led to reproducible estimates of each homolog's amino-acid preferences.
{\bf(A)} For each replicate of each homolog, we deep sequenced the starting plasmid libraries (DNA) and the selected virus libraries (virus).
Bars show the per-codon mutation frequency averaged across all sites after subtracting mutation frequencies determined using wildtype controls (Fig \ref{suppfig:ch3_mutfreqs_supp}).
Mutations leading to stop codons were purged during the selection step to 3-17\% their starting frequencies (see red numbers), indicating strong purifying selection.
Nonsynonymous mutations decreased to 40-50\% their starting frequencies, consistent with a large fraction of amino-acid mutations being deleterious to Env.
Synonymous mutations only decreased to 87-95\% their starting frequencies, as would be expected if synonymous mutations tend to have more neutral effects than nonsynonymous changes.
Panels {\bf(B)} and {\bf(C)} show the correlation in our estimates of each homolog's site-specific amino-acid preferences between experimental replicates, with {\bf(B)} showing the correlation between BF520 replicates and {\bf(C)} showing the correlation between BG505 replicates.
Each correlation plot reports the associated Pearson correlation coefficient, which ranged from 0.59-0.75, indicating that our estimates are largely repeatable between replicates.
}
%\figsupp{test text}{\includegraphics[width=5cm]{figures/mutfreqs_supp/mutfreqs_supp}}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=5in]{figures/mutfreqs_supp/mutfreqs_supp}}
\caption{\label{suppfig:mutfreqs_supp}
{\bf Per-codon mutation frequencies of the mutant libraries and wildtype controls before and after selection.}
This figure is similar to Fig \ref{fig:mutfreqs}, but it reports \textit{non}-error-corrected per-codon mutation frequencies for each of the mutant libraries and wildtype controls before and after selection for {\bf (A)} BG505 and {\bf (B)} BF520.
Specifically, for each each replicate, wt-DNA and mut-DNA refer to the pre-selection wildtype and mutant plasmids, respectively; and wt-virus and mut-virus refer to the post-selection wildtype and mutant viruses, respectively.
Note, there is only a single replicate of the wt-DNA sample for BF520.
The bars show frequencies of different codon-level mutations.
The left bar for each sample groups mutations as being either synonymous, non-synonymous, or leading to the introduction of a stop codon.
The right bar for each sample groups mutations as introducing a single-nucleotide codon mutation (e.g., aaa $\rightarrow$ aaT), double-nucleotide codon mutation (e.g., aaa $\rightarrow$ aTT), or triple-nucleotide mutation (e.g., aaa $\rightarrow$ TTT).
Each wildtype plasmid shows a low rate of single-nucleotide codon mutations, consistent with errors from PCR and deep sequencing.
Importantly, the wildtype plasmids have a much lower mutation rate than the mutant plasmids.
The mutation frequency of single-nucleotide mutations substantially increases for each wildtype virus, which is expected due to errors from viral replication.
Comparing the wildtype viruses to the mutant viruses suggests that roughly a third to a half of mutations in the mutant viruses arise due to to errors from PCR, deep sequencing, and viral replication, underlying the importance of estimating these error rates using the controls.
When inferring Env's amino-acid preferences, we use the wildtype controls to statistically correct for such errors, as described in the Methods.
}
\end{figure}

>>>>

\begin{figure}
\centerline{\includegraphics[width=5in]{figures/tree/tree}}
\caption{\label{fig:tree}
Phylogenetic tree showing the relationship among BG505 and BF520.
{\bf(A)} A tree of group-M sequences with 20 sequences per subtype for subtypes A, B, C, D, F, and G.
{\bf(B)} A tree of subtype-A sequences with 120 sequences total, rooted using a subtype-B sequence as an outgroup.
BG505 and BF520 cluster within subtype A, while LAI clusters within subtype B.
The alignments we used to make these trees are the same alignments we used in the below \texttt{phydms} analysis, excluding the outgroup sequence used to root the subtype-A tree.
The scale bars are in units of codon substitutions per site.
}
\figdata{The alignments used to make the group-M and subtype-A trees are in files \ref{suppfile:grpM} and \ref{suppfile:cladeA}, respectively}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=5in]{figures/mutfreqs/mutfreqs}}
\caption{\label{fig:mutfreqs}
The deep mutational scanning experiments imposed strong purifying selection and led to reproducible estimates of each homolog's amino-acid preferences.
{\bf(A)} For each replicate of each homolog, we deep sequenced the starting plasmid libraries (DNA) and the selected virus libraries (virus).
Bars show the per-codon mutation frequency averaged across all sites after subtracting mutation frequencies determined using wildtype controls (Fig \ref{suppfig:ch3_mutfreqs_supp}).
Mutations leading to stop codons were purged during the selection step to 3-17\% their starting frequencies (see red numbers), indicating strong purifying selection.
Nonsynonymous mutations decreased to 40-50\% their starting frequencies, consistent with a large fraction of amino-acid mutations being deleterious to Env.
Synonymous mutations only decreased to 87-95\% their starting frequencies, as would be expected if synonymous mutations tend to have more neutral effects than nonsynonymous changes.
Panels {\bf(B)} and {\bf(C)} show the correlation in our estimates of each homolog's site-specific amino-acid preferences between experimental replicates, with {\bf(B)} showing the correlation between BF520 replicates and {\bf(C)} showing the correlation between BG505 replicates.
Each correlation plot reports the associated Pearson correlation coefficient, which ranged from 0.59-0.75, indicating that our estimates are largely repeatable between replicates.
}
\figsupp[Per-codon mutation frequencies of the mutant libraries and wildtype controls before and after selection.\label{mutfreqs_supp}]{Per-codon mutation frequencies of the mutant libraries and wildtype controls before and after selection.
This figure is similar to Fig \ref{fig:mutfreqs}, but it reports \textit{non}-error-corrected per-codon mutation frequencies for each of the mutant libraries and wildtype controls before and after selection for {\bf (A)} BG505 and {\bf (B)} BF520.
Specifically, for each each replicate, wt-DNA and mut-DNA refer to the pre-selection wildtype and mutant plasmids, respectively; and wt-virus and mut-virus refer to the post-selection wildtype and mutant viruses, respectively.
Note, there is only a single replicate of the wt-DNA sample for BF520.
The bars show frequencies of different codon-level mutations.
The left bar for each sample groups mutations as being either synonymous, non-synonymous, or leading to the introduction of a stop codon.
The right bar for each sample groups mutations as introducing a single-nucleotide codon mutation (e.g., aaa $\rightarrow$ aaT), double-nucleotide codon mutation (e.g., aaa $\rightarrow$ aTT), or triple-nucleotide mutation (e.g., aaa $\rightarrow$ TTT).
Each wildtype plasmid shows a low rate of single-nucleotide codon mutations, consistent with errors from PCR and deep sequencing.
Importantly, the wildtype plasmids have a much lower mutation rate than the mutant plasmids.
The mutation frequency of single-nucleotide mutations substantially increases for each wildtype virus, which is expected due to errors from viral replication.
Comparing the wildtype viruses to the mutant viruses suggests that roughly a third to a half of mutations in the mutant viruses arise due to to errors from PCR, deep sequencing, and viral replication, underlying the importance of estimating these error rates using the controls.
When inferring Env's amino-acid preferences, we use the wildtype controls to statistically correct for such errors, as described in the Methods.}{\includegraphics[width=5in]{figures/mutfreqs_supp/mutfreqs_supp}}
\end{figure}

@@
